\section*{Problem 3}

The goal of this problem is to exploit the power of the recommended tools rather than elaborating the questions by hand

\begin{description}
  \item[(a)] In mathematics, a \emph{group} is defined to be a set $G$ with an element $I \in G$, a binary operator $\ast$ and a unary operator \emph{inv} satisfying
      \begin{center}
      $ x \ast(y \ast z)=(x\ast y) \ast z, x \ast I = x $ and $ x \ast inv(x) = I, $
      \end{center}
      for all $x, y, z \in G$. Determine whether in every group each of the four properties
      \begin{center}
      $I \ast x = x, inv(inv(x)) = x, inv(x) \ast x = I $ and $ x \ast y = y \ast x $
      \end{center}
      holds for all $x, y \in G$. If a property does not hold, determine the size of the smallest finite group for which it does not hold.
  \item[(b)] A term rewrite system consists of the single rule
  \[ a(x, a(y, a(z, u))) \rightarrow a(y, a(z, a(x, u)))), \]

  in which $a$ is a binary symbol and $x, y, z, u$ are variables. Moreover, there are constants $b, c, d, e, f, g$. Determine whether $c$ and $d$ may be swapped in $a(b, a(c, a(d, a(e, a(f, a(b, g))))))$ by rewriting, that is, $a(b, a(c, a(d, a(e, a(f, a(b, g))))))$ rewrites in a finite number of steps to $a(b, a(d, a(c, a(e, a(f, a(b, g))))))$.

\end{description}

\vspace{4mm}

\subsection*{Solution:}

\begin{description}
  \item[(a)] In this problem, three assumptions are given. So we use {\tt Prover9} to prove the four properties straightforwardly. The expressions in {\tt Prover9} are as follows. Here we denote \emph{inv(x)} as $x'$ for the sake of simplicity.

\vspace{3mm}

{\footnotesize

{\tt formulas(assumptions).}

{\tt \% Group definition}

{\tt x * I = x.}

{\tt x * x' = I.}

{\tt x * (y * z) = (x * y) * z.}

{\tt end\_of\_list.}

{\tt formulas(goals).}

{\tt I * x = x.}

{\tt x'' = x.}

{\tt x' * x = I.}

{\tt x * y = y * x.}

{\tt end\_of\_list.}

}

\vspace{2mm}

  After applying {\tt prover9 -f part2\_3a.in}, we found that the first 3 properties are proved, but the fourth one, which implies the property of commutativity, is failed. In order to determine the size of the smallest finite group for which it does not hold, we apply {\tt mace4 part$2\_3a$.in} to find the smallest noncommutative group by finding the counterexample to the statement that all groups are commutative. {\tt Mace4} exits with 1 model. The model is as follows.

\vspace{3mm}

{\tt ============================== DOMAIN SIZE 6 =========================}

{\tt ============================== MODEL =================================}

{\tt interpretation( 6, [number=1, seconds=0], [}

{\tt \ \ \ \ \ \ \ \ function(I, [ 0 ]),}

{\tt \ \ \ \ \ \ \ \ function(c1, [ 1 ]),}

{\tt \ \ \ \ \ \ \ \ function(c2, [ 2 ]),}

{\tt \ \ \ \ \ \ \ \ function('(\_), [ 0, 1, 2, 4, 3, 5 ]),}

{\tt \ \ \ \ \ \ \ \ function(*(\_,\_), [}

{\tt \	\	\	\ \ \ \	\	\	\ \ \ \ \ \ 0, 1, 2, 3, 4, 5,}

{\tt \	\	\	\ \ \ \	\	\	\ \ \ \ \ \ 1, 0, 3, 2, 5, 4,}

{\tt \	\	\	\ \ \ \	\	\	\ \ \ \ \ \ 2, 4, 0, 5, 1, 3,}
	
{\tt \	\	\	\ \ \ \	\	\	\ \ \ \ \ \ 3, 5, 1, 4, 0, 2,}

{\tt \	\	\	\ \ \ \	\	\	\ \ \ \ \ \ 4, 2, 5, 0, 3, 1,}

{\tt \	\	\	\ \ \ \	\	\	\ \ \ \ \ \ 5, 3, 4, 1, 2, 0 ])}

{\tt ]).}

{\tt ============================== end of model ==========================}

\vspace{3mm}

Therefore, {\tt Mace4} proved that the size of the smallest finite group, for which $ x \ast y = y \ast x $ does not hold, is 6.

  \item[(b)] To determine the possibility of the rewriting in a finite number of steps, we use {\tt Prover9} to prove the possibility.

  This problem gives only a single rewriting rule, so the TRS $R$ consists of the rule  \[ a(x, a(y, a(z, u))) \rightarrow_{R} a(y, a(z, a(x, u)))), \]
  %Here $R$ means single rewrite step.

  The rewriting terms are the closed terms composed from the constants $b, c, d, e, f, g$. We want to show that the rewriting of \[ a(b, a(c, a(d, a(e, a(f, a(b, g)))))) \rightarrow a(b, a(d, a(c, a(e, a(f, a(b, g)))))). \] can be done in zero or more rewrite steps.
  %Here $RR$ means zero or more rewrite steps.


      %\[ R(a(x, u), a(u, x)). \]
      %\[ R(x, y) \rightarrow R(a(x, z), a(y, z)). \]
      %\[ R(x, y) \rightarrow R(a(z, x), a(z, y)).\]

      Expressed in {\tt Prover9}:

\vspace{2mm}

{\footnotesize

{\tt formulas(assumptions).}

%{\tt R(a(x, u), a(u, x)).}

%{\tt \% the given single rule}

{\tt R(a(x,a(y,a(z,u))),a(y,a(z,a(x,u)))).}

{\tt R(x,y) -> R(a(x,z),a(y,z)).}

{\tt R(x,y) -> R(a(z,x),a(z,y)).}

{\tt RR(x,x). }

{\tt (RR(x,y) \& R(y,z)) -> RR(x,z). }

{\tt end\_of\_list.}

{\tt formulas(goals).}

{\tt RR(a(b,a(c,a(d,a(e,a(f,a(b,g)))))),a(b,a(d,a(c,a(e,a(f,a(b, g))))))).}

{\tt end\_of\_list.  }

}

\vspace{3mm}

$a$ = $a$ binary operator, $R$ = single rewrite step, and $RR$ = zero or more rewrite steps.

  After running {\tt prover9 -f  part2\_3b.in}, {\tt Prover9} exits with 1 proof. Therefore, $c$ and $d$ can be swapped in $a(b, a(c, a(d, a(e, a(f, a(b, g))))))$ by rewriting in a finite number of steps to $a(b, a(d, a(c, a(e, a(f, a(b, g))))))$. The proof is in the following.
 
\vspace{3mm} 

{\footnotesize

{\tt ============================== PROOF =================================}

{\tt \% Proof 1 at 0.09 (+ 0.03) seconds.}

{\tt \% Length of proof is 22.}

{\tt \% Level of proof is 7.}

{\tt \% Maximum clause weight is 27.}

{\tt \% Given clauses 195.}

{\tt 2 R(x,y) -> R(a(z,x),a(z,y)) \# label(non\_clause).  [assumption].}

{\tt 3 RR(x,y) \& R(y,z) -> RR(x,z) \# label(non\_clause).  [assumption].}

{\tt 4 RR(a(b,a(c,a(d,a(e,a(f,a(b,g)))))),a(b,a(d,a(c,a(e,a(f,a(b,g))))))) \#
label(non\_clause) \# label(goal).}

$\cdots \cdots \cdots$

{\tt 800 -RR(a(b,a(c,a(d,a(e,a(f,a(b,g)))))),a(c,a(d,a(f,a(e,a(b,a(b,g))))))).  [ur(9,b,38,a,c,116,a)].}

{\tt 1604 RR(a(x,a(y,a(z,a(u,a(w,a(v5,v6)))))),a(y,a(z,a(w,a(u,a(v5,a(x,v6))))))).  [ur(9,a,319,a,b,170,a)].}

{\tt 1605 \$F.  [resolve(1604,a,800,a)].}

{\tt ============================== end of proof ==========================}

}

\vspace{3mm} 

\end{description}

\subsection*{Remark:}

Since Problem 3$a$ gives the rewriting rules and the goals explicitly, its expression in {\tt Prover9} also goes straightforwardly. The searching for a finite model with the last property, which is failed in the proof of {\tt Prover9}, is done by the tool {\tt Mace4}. Problem 3$b$ requires the definition of the term rewrite system (TRS) $R$ to complete the proof. {\tt Prover9} is an automated theorem prover for first-order and equational logic, and {\tt Mace4} is an automated searcher for finite models and counterexamples. Both the tools are very powerful in automated reasoning. Thanks to their programmer, William McCune.

%\subsection*{Generalization:}




