\documentclass[11pt]{article}
\usepackage{a4wide}
\usepackage{latexsym}
\usepackage{amssymb}
\usepackage{epic}
\usepackage{graphicx}
%\pagestyle{empty}
\newcommand{\tr}{\mbox{\sf true}}
\newcommand{\fa}{\mbox{\sf false}}
\newcommand{\bimp}{\leftrightarrow}


\begin{document}

\section*{
\begin{center}
Practical Assignment Part 1\\
Automated Reasoning 2IMF25\\
\end{center}
}

\begin{center}
Technische Universiteit Eindhoven\\
Jiahuan Zhang 0896785 (j.4.zhang@student.tue.nl)\\
Hector Joao Rivera Verduzco 0977393 (h.j.rivera.verduzco@student.tue.nl)\\

\end{center}
\date{\today}
\vspace{5mm}

\section*{Problem 1}

Six trucks have to deliver pallets of obscure building blocks to a magic factory. Every truck has a capacity of 7800 kg and can carry at most eight pallets. In total, the following has to be delivered:
\begin{itemize}
  \item Four pallets of nuzzles, each of weight 700 kg.
  \item A number of pallets of prittles, each of weight 800 kg.
  \item Eight pallets of skipples, each of weight 1000 kg.
  \item Ten pallets of crottles, each of weight 1500 kg.
  \item Five pallets of dupples, each of weight 100 kg.
\end{itemize}
Prittles and crottles are an explosive combination: they are not allowed to be put in the same truck.\\
Skipples need to be cooled; only two of the six trucks have facility for cooling skipples.\\
Dupples are very valuable; to distribute the risk of loss no two pallets of dupples may be in the same truck.\\
Investigate what is the maximum number of pallets of prittles that can be delivered, and show how for that number all pallets may be divided over the six trucks.

\vspace{4mm}

\subsection*{Solution:}

We generalize the problem to offer $n$ trucks to deliver the pallets. $n$ is a positive integer. Then we introduce some variables $t_{ij}$ for $i = 1, ..., n$ and $j = 1,...,5$, which represents the number of pallets of obscure building blocks $j$ on the $i$-th truck. $t_{ij}$ is a natural number. Every truck has a capacity, which is denoted as $C$ kg, and the maximum number of pallets each truck can carry is $M$.

The following table shows which index is related to which building material:

\begin{center}
\begin{tabular}{|c|c|c|c|c|c|}
  \hline
  % after \\: \hline or \cline{col1-col2} \cline{col3-col4} ...
    & nuzzles & prittles & skipples & crottles & dupples \\
  $j$ & 1 & 2 & 3 & 4 & 5 \\
  \hline
\end{tabular}
\end{center}

The pallets of the same building blocks have the same weight, $weight(j)$. Each kind of building blocks has a predefined number of pallets, $\sharp pallets(j)$.

Now we consider the conditions for the delivery.

\begin{description}
  \item[Clause 1:] $t_{ij}$ should be no less than $0$.

  This is expressed by the formula
  \[ \bigwedge_{i,j:1 \leq i \leq n \wedge 1 \leq j \leq 5} t_{ij} \geq 0.\]
  \item[Clause 2:] Every truck has a maximum capacity of $C$ kg.

  \[ \bigwedge_{i=1}^n  (\sum_{j=1}^5 t_{ij} \times weight(j)) \leq C .\]
  \item[Clause 3:] Every truck can carry at most $M$ pallets.

  \[ \bigwedge_{i=1}^n (\sum_{j=1}^5 t_{ij}) \leq M .\]
  \item[Clause 4:] The total numbers of the pallets of each obscure building blocks should be exactly the same as the given number by $\sharp pallets(j)$.

  \[ \bigwedge_{j=1}^5 (\sum_{i=1}^n t_{ij}) = \sharp pallets(j) .\]

  \item[Clause 5:] Prittles and crottles are not allowed to be put in the same truck. So in every truck the number of prittles or the number of crottles should be zero.

  \[ \bigwedge_{i=1}^n t_{i2}=0 \vee t_{i4}=0 .\]
  \item[Clause 6:] Only two of the trucks can deliver skipples.

      So let's assume the $n$-th and $n-2$-th trucks can deliver skipples. Then the amount of skipples in these two trucks should be equal to the total amount of this material.

  \[ t_{n3} + t_{(n-2)3} = \sharp pallets(3) .\]
  \item[Clause 7:] No two pallets of dupples may be in the same truck. This means that at most one pallet of dupple is allowed in each truck.

  \[ \bigwedge_{i=1}^n t_{i5} \leq 1 .\]
\end{description}

The total formula now consists of the conjunction of all these
ingredients, that is,
\[ \bigwedge_{i,j:1 \leq i \leq n, 1 \leq j \leq 5} t_{ij} \geq 0 \;\; \wedge \]
\[ \bigwedge_{i=1}^n  (\sum_{j=1}^5 t_{ij} \times weight(j)) \leq C \;\; \wedge \]
\[ \bigwedge_{i=1}^n (\sum_{j=1}^5 t_{ij}) \leq M \;\; \wedge \]
\[ \bigwedge_{j=1}^5 (\sum_{i=1}^n t_{ij}) = \sharp pallets(j) \;\; \wedge \]
\[ \bigwedge_{i=1}^n t_{i2}=0 \vee t_{i4}=0 \;\; \wedge \]
\[ t_{n3} + t_{(n-2)3} = \sharp pallets(3) \;\; \wedge \]
\[ \bigwedge_{i=1}^n t_{i5} \leq 1 \]

This formula is easily expressed in SMT syntax. For Problem 1, we define the number of trucks $n = 6$, the maximum number of pallets in each truck $M = 8$, the capacity of the trucks as $C = 7800$ kg, and the parameters $weight(j)$ and $\sharp pallets(j)$ as specified in the table below.
\begin{center}
\begin{tabular}{|c|c|c|c|c|c|}
  \hline
  % after \\: \hline or \cline{col1-col2} \cline{col3-col4} ...
    & nuzzles & prittles & skipples & crottles & dupples \\
  $j$ & 1 & 2 & 3 & 4 & 5 \\
  $weight(j)$ & 700kg & 800kg & 1000kg & 1500kg & 100kg \\
  $\sharp pallets(j)$ & 4 & $P$ & 8 & 10 & 5 \\
  \hline
\end{tabular}
\end{center}

With the values above, we get the following Yices codes.

{\footnotesize

{\tt benchmark Part1\_1.smt}

{\tt :logic $QF\_ALIA$}

{\tt :extrafuns (}

{\tt (t11 Int) (t12 Int) (t13 Int) (t14 Int) (t15 Int) }

{\tt (t21 Int) (t22 Int) (t23 Int) (t24 Int) (t25 Int) }

$\cdots \cdots$

{\tt (t61 Int) (t62 Int) (t63 Int) (t64 Int) (t65 Int) }

{\tt )}

{\tt :formula}

{\tt   (and}

{\tt (>= t11 0) (>= t12 0) (>= t13 0) (>= t14 0) (>= t15 0)}

{\tt (>= t21 0) (>= t22 0) (>= t23 0) (>= t24 0) (>= t25 0)}

$\cdots \cdots$

{\tt (>= t61 0) (>= t62 0) (>= t63 0) (>= t64 0) (>= t65 0) }

{\tt (<= (+ (* t11 700) (* t12 800) (* t13 1000) (* t14 1500) (* t15 100)) 7800)}

{\tt (<= (+ (* t21 700) (* t22 800) (* t23 1000) (* t24 1500) (* t25 100)) 7800)}

$\cdots \cdots$

{\tt (<= (+ (* t61 700) (* t62 800) (* t63 1000) (* t64 1500) (* t65 100)) 7800) }

{\tt (<= (+ t11 t12 t13 t14 t15) 8)}

{\tt (<= (+ t21 t22 t23 t24 t25) 8)}

$\cdots \cdots$

{\tt (<= (+ t61 t62 t63 t64 t65) 8)}

{\tt (= (+ t11 t21 t31 t41 t51 t61) 4)}

{\tt (= (+ t12 t22 t32 t42 t52 t62) 18)}

{\tt (= (+ t13 t23 t33 t43 t53 t63) 8)}

{\tt (= (+ t14 t24 t34 t44 t54 t64) 10)}

{\tt (= (+ t15 t25 t35 t45 t55 t65) 5)}

{\tt (or (= t12 0) (= t14 0)) }

{\tt (or (= t22 0) (= t24 0)) }

$\cdots \cdots$

{\tt (or (= t62 0) (= t64 0)) }

{\tt (= (+ t43 t63) 8) }

{\tt (<= t15 1) (<= t25 1) (<= t35 1) (<= t45 1) (<= t55 1) (<= t65 1)}

{\tt ) }
}

Applying {\tt yices-smt -m Part1\_1.smt} several times, we find when the number of pallets of prittles $P$ is 19, it is UNSAT. When the number is 18, it is SAT. Therefore, we conclude that the maximal number of pallets of prittles is 18.
The following result is yielded within a fraction of a second:

{\footnotesize

{\tt sat}

{\tt (= t52 7)}

{\tt (= t11 0)}

{\tt (= t14 5)}

{\tt (= t34 0)}

{\tt (= t21 4)}

{\tt (= t15 1)}

{\tt (= t44 2)}

{\tt (= t45 1)}

{\tt (= t54 0)}

{\tt (= t33 0)}

{\tt (= t35 0)}

{\tt (= t51 0)}

{\tt (= t23 0)}

{\tt (= t53 0)}

{\tt (= t65 1)}

{\tt (= t13 0)}

{\tt (= t62 0)}

{\tt (= t25 1)}

{\tt (= t32 8)}

{\tt (= t41 0)}

{\tt (= t63 4)}

{\tt (= t43 4)}

{\tt (= t42 0)}

{\tt (= t12 0)}

{\tt (= t64 0)}

{\tt (= t55 1)}

{\tt (= t22 0)}

{\tt (= t24 3)}

{\tt (= t61 0)}

{\tt (= t31 0)}

}
\vspace{3mm}
The following table shows the solution for this specific problem.

\vspace{2mm}
  \begin{center}
  \begin{tabular}{|l|c|c|c|c|c|}
    \hline
    % after \\: \hline or \cline{col1-col2} \cline{col3-col4} ...
     & nuzzles & prittles & skipples & crottles & dupples \\
    Truck 1 & 0 & 0 & 0 & 5 & 1 \\
    Truck 2 & 4 & 0 & 0 & 3 & 1 \\
    Truck 3 & 0 & 8 & 0 & 0 & 0 \\
    Truck 4 & 0 & 0 & 4 & 2 & 1 \\
    Truck 5 & 0 & 7 & 0 & 0 & 1 \\
    Truck 6 & 0 & 3 & 4 & 0 & 1 \\
    \hline
  \end{tabular}
\end{center}
\subsection*{Remark}
In this problem, we are required to find out the maximum number of the pallets of prittles. Such kind of maximum values searching is likely to become very time-consuming since there is no explicitly close upper bound to start the searching.

For this problem, we first find out the theoretically possible maximum number of the pallets of prittles through calculation. Because the capacity of each track is given, and it is also provided the total weights of the pallets of the other obscure building blocks, we can estimate the number.

  $\frac{7800 \times 6 - 700\times4 - 1000\times8 - 1500\times10 - 100\times5}{800} = 25.625$

Because the number of pallets is a natural number, the maximum number should not be more than 25. Therefore, we can start the debugging from $P = 25$ downwards.

\subsection*{Generalization}

We generalized all the invariants except the total number of kinds of the building blocks in this problem, because most of the requirements come from the features of different building blocks, which indicates that adding more different building blocks will generate more requirements, and as a consequent, the entire formula we summarized in the end of the solution is not applicable.

One more note is for the people who are interested in setting the number of trucks, $n$, larger than $10$. They need to take care about the notations of the variables $t_{ij}$. For instance, the number of pallets of building blocks labeled $1$ on the eleventh truck, $t_{11,1}$, is expressed as  $t111$ in Yices codes. This expression can also represent the number of pallets of building blocks labeled $11$ on the first truck, $t_{1,11}$. An extra symbol between the two numbers $i$ and $j$ is required to avoid this ambiguity.

Decreasing the number of trucks is more of interest. We did some testing to figure out how many pallets of prittles can satisfy these conditions with fewer trucks. When there are five trucks in all, ten pallets of prittles can be delivered with the satisfiability of the conditions. With fewer trucks, no satisfiability is reached. Since one truck can only deliver one pallet of drupples due to the condition, the number of trucks has to be larger than the number of pallets of dupples for delivery. It can be expressed as this
\[ \sharp pallets(5) \leq m \]

\newpage

\input{Problem2.tex}

\newpage
\input{Problem3.tex}

\newpage
\input{Problem4.tex}

\end{document}
